# Intro to Data Science - HW 6
##### Copyright Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva


```{r}
# Enter your name here: Nora Lin
```


### Attribution statement: (choose only one and delete the rest)


```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```

Last assignment we explored **data visualization** in R using the **ggplot2** package. This homework continues to use ggplot, but this time, with maps.  In addition, we will merge datasets using the built-in **merge( )** function, which provides a similar capability to a **JOIN in SQL** (don't worry if you do not know SQL). Many analytical strategies require joining data from different sources based on a **“key”** – a field that two datasets have in common. 

## Step 1: Load the population data

A.	Read the following JSON file, https://intro-datascience.s3.us-east-2.amazonaws.com/cities.json and store it in a variable called **pop**.

Examine the resulting pop dataframe and add comments explaining what each column contains.

```{r}
library(jsonlite)
dataset <- url("https://intro-datascience.s3.us-east-2.amazonaws.com/cities.json")
readlines <- jsonlite::fromJSON(dataset)
pop<- readlines
```

B.	Calculate the **average population** in the dataframe. Why is using mean() directly not working? Find a way to correct the data type of this variable so you can calculate the average (and then calculate the average)

Hint: use **str(pop)** or **glimpse(pop)** to help understand the dataframe


```{r}
str(pop)
pop$population <- as.numeric(pop$population)
mean(pop$population)
#The mean() is not directly working because the structure of the column population is a character. The average population is 131132.4
```

C.	What is the population of the smallest city in the dataframe? Which state is it in?
```{r}
pop[pop$population==min(pop$population),]
#The population of the smallest city is Panama City with a population fo 36877
```

## Step 2: Merge the population data with the state name data

D)	Read in the state name .csv file from the URL below into a dataframe named **abbr** (for "abbreviation") – make sure to use the read_csv() function from the tidyverse package: <br>
https://intro-datascience.s3.us-east-2.amazonaws.com/statesInfo.csv

```{r}
library(tidyverse)
abbr <- read_csv("https://intro-datascience.s3.us-east-2.amazonaws.com/statesInfo.csv")

```

E)	To successfully merge the dataframe **pop** with the **abbr** dataframe, we need to identify a **column they have in common** which will serve as the **“key”** to merge on. One column both dataframes have is the **state column**. The only problem is the slight column name discrepancy – in **pop**, the column is called **“state”** and in **abbr** – **“State.”** These names need to be reconciled for the merge() function to work. Find a way to rename **abbr’s “State”** to **match** the **state column in pop**.    


```{r}
#str(abbr)
colnames(abbr)[1] <- "state"
```

F)	Merge the two dataframes (using the **‘state’ column** from both dataframes), storing the resulting dataframe in **dfNew**.


```{r}
dfNew <- merge(pop,abbr,by.x="state",by.y="state")
dfNew
```

G)	Review the structure of **dfNew** and explain the columns (aka attributes) in that dataframe.


```{r}
str(dfNew)
```

## Step 3: Visualize the data

H)	Plot points (on top of a map of the US) for **each city**. Have the **color** represent the **population**.


```{r}
library(ggplot2)
library(maps)
```

```{r}
ggplot()+
  geom_map(data=us,map=us,aes(map_id = region, x=long, y=lat),color="black",fill="white")+
  geom_point(data=dfNew,aes(longitude,latitude,group=city,color=population))
```

I)	Add a block comment that criticizes the resulting map. It’s not very good.


```{r}
#This map isn't great.  It looks squished. 
```

## Step 4: Group by State


J)	Use group_by and summarise to make a dataframe of state-by-state population. Store the result in **dfSimple**.


```{r}
dfSimple <-dfNew %>%
  group_by(state)%>%
  summarise(total_population= sum(population))
dfSimple
```

K)	Name the most and least populous states in **dfSimple** and show the code you used to determine them.


```{r}
#Least populous states: 
dfSimple[dfSimple$total_population==min(dfSimple$total_population)]

#The least populous states are alaska, district  of columbia, hawaii, maine and vermont.
```
```{r}
#most populous states:
dfSimple[dfSimple$population==max(dfSimple$population),]
# the most populous state is california
```

## Step 5: Create a map of the U.S.,  with the color of the state representing the state population

L) Make sure to expand the limits correctly and that you have used **coord_map** appropriately.

```{r}
dfNew2 <- merge(dfNew, dfSimple, by.x= "state", by.y="state")
dfNew2
```

```{r}
ggplot(dfNew2)+
  aes(longitude,latitude,group=state)+
  geom_polygon(aes(fill=total_population))
```




